Fast gravitational wave parameter estimation without compromises

### 2023-07-21 src/tex/ms.tex
Add why flow is important

% GPU helps packing a shit ton of waveform evaluation
Another benefit of accelerators is the parallel evaluation of waveforms. Since
the waveform model we use can be evaluated at any given time or frequency
independently, this means computing a waveform can be trivially parallelized
over frequency bins. On an \textsc{Nvidia} A100 GPU, we can evaluate the
waveform model $\sim \mathcal{O}(10^9)$ times in a second for different
frequencies or source parameters. Together with the heterodyned likelihood, this
allows us to run thousands of parallel chains in a PE run. The high throughput
of waveform evaluations unlocks the potential of the \textsc{flowMC} sampling
algorithm.

>>>>

% GPU helps packing a shit ton of waveform evaluation
Another benefit of accelerators is the parallel evaluation of waveforms. Since
the waveform model we use can be evaluated at any given time or frequency
independently, this means computing a waveform can be trivially parallelized
over frequency bins. On an \textsc{Nvidia} A100 GPU, we can evaluate the
waveform model $\sim \mathcal{O}(10^9)$ in a second for different
frequencies or source parameters. Together with the heterodyned likelihood, this
allows us to run thousands of parallel chains in a PE run. The high throughput
of waveform evaluations unlocks the potential of the \textsc{flowMC} sampling
algorithm.

\corr{The interplay between \textsc{flowMC} and GPU acceleration is critical to
the performance of the code. A standard PE run using a nested sampler or
standard MCMC requires around $10^6$ waveform evaluations, Assuming $10^5$
frequency bins per waveform, that corresponds to evaluating the waveform model
at $10^11$ frequency-parameter points. One would think this can be achieved
within $\sim 100\textrm{s}$ by a A100 GPU given the benchmark mentioned above.
However, a sampler requires sequential evaluation to accept or reject a sample,
where what a GPU provide is parallel evaluation of the waveform, i.e. it can run
many chains in parallel, but not many steps with the same amount of computation.
This means one cannot just obtain the posterior samples by running the waveform
model on a GPU with a typical sampler, in the sense that we cannot run many
chains for one step and expect unbias posterior samples.

This is precisely why \textsc{flowMC} is necessary for our approach. Since every
accepted sample generated using the normalizing flow as proposal is independent,
and the sampling from the normalizing flow is an embarrassingly parallel
process, the name of the game then becomes training the normalizing flow as fast as
possible. Having many chains helps to produce the necessary amount of samples
for training the normalizing flow, which means now we can meaningfully harvest
the benefit of having large throughput of GPU. Without either of this two,
our approach would be significantly less performant.}
